Proving Logical Atomicity using Lock Invariants\thanks{Supported by the National Science Foundation via grant \#1811894.}

### 2022-05-18 concurrency/ASL 2022/intro.tex
edits and space hacks

In this paper, we adapt this \LB{This reads like ``we carry out Zhang's approach in VST''}\wm{That's basically true.} approach to the Verified Software Toolchain~\cite{plfcc}, a system for proving correctness of C programs that has recently been extended with support for general ghost state and atomicity proofs. The soundness proof for VST~\cite{cpm} lifts the single-threaded correctness theorem of the CompCert compiler~\cite{compcert} to a concurrent setting, and relies explicitly on style 1 lock specs; modifying this proof to rely on atomic operations instead is both theoretically and practically daunting. We demonstrate that 1) \textbf{our approach can be used to prove atomic specs for interesting data structures with fine-grained locking, starting from style 1 lock specs}, and 2) \textbf{these proofs require some additional complications that can be avoided with style 2 lock specs}. The key difference is that in style 1 a thread must gain ownership of a handle to a component's lock before acquiring the lock and reading/modifying the component, while in style 2 the locks and components alike can be considered part of a single abstract state that is accessed at each lock access. However, the top-level specifications proved for the data structure are not affected by the complexity of the lock specs, so proofs of clients of the data structure are no more complex than they would be otherwise.

@@

\LB{In contrast to all previous wroks, we our work is on full C implementations of binary search trees rather than pseudo code. This is enabled by recently added incorporation of (slightly limited) Iris-style ghost state and invariants (and Iris prof mode), and of TaDa-style atomic specs in VST (citation to William's technical report?)}\wm{It's currently a rejected ITP paper rather than a technical report, sadly.}

>>>>

In this paper, we adapt this approach to the Verified Software Toolchain~\cite{plfcc}, a system for proving correctness of C programs that has recently been extended with support for general ghost state and atomicity proofs. The soundness proof for VST~\cite{cpm} lifts the single-threaded correctness theorem of the CompCert compiler~\cite{compcert} to a concurrent setting, and relies explicitly on style 1 lock specs; modifying this proof to rely on atomic operations instead is both theoretically and practically daunting. We demonstrate that 1) \textbf{our approach can be used to prove atomic specs for interesting data structures with fine-grained locking, starting from style 1 lock specs}, and 2) \textbf{these proofs require some additional complications that can be avoided with style 2 lock specs}. The key difference is that in style 1 a thread must gain ownership of a handle to a component's lock before acquiring the lock and reading/modifying the component, while in style 2 the locks and components alike can be considered part of a single abstract state that is accessed at each lock access. However, the top-level specifications proved for the data structure are not affected by the complexity of the lock specs, so proofs of clients of the data structure are no more complex than they would be otherwise.

@@
